---
title: "stl_lead_inequity.qmd"
author: "OP"
format: html
editor: visual
execute: 
  echo: false
  warning: false
---

```{r}
library(tidyverse)
library(janitor)

```

```{r}
stl_lead <- read_csv(here::here("data", "stl_lead.csv"), na = "#NA") %>% 
  clean_names()
  #rename(cl_a = active_chl_a_ug_l)
```

-   Do some basic exploration of the dataset (e.g. using summary, data visualizations and summary statistics).

```{r}

mean(stl_lead$poverty_tot)

```

```{r}
sd(stl_lead$poverty_tot)
```

```{r}
head(stl_lead$poverty_tot)
```

```{r}
tail(stl_lead)
```

```{r}
ggplot(data = stl_lead, aes (x = total_pop, y = stl_lead$poverty_tot))+
  geom_point()
```

-   In a new code chunk, from `stl_lead` create a new data frame called `stl_lead_prop` that has one additional column called `prop_white` that returns the percent of each census tract identifying as white (variable `white` in the dataset divided by variable `totalPop`, times 100). You may need to do some Googling. Hint: `dplyr::mutate(new_col = col_a / col_b)`will create a new column `new_col` that contains the value of `col_a / col_b`

```{r}
stl_lead_prop <- data.frame(stl_lead) %>% 
  dplyr::mutate(prop_white = (white/total_pop)*100 )
```

### **Create a scatterplot**

-   In a new code chunk, create a scatterplot graph of the percentage of children in each census tract with elevated blood lead levels (`pctElevated`) versus the percent of each census tract identifying as white.

-   Customize by updating several aesthetics (e.g. size, opacity (see `alpha =`), color, etc.)

-   Store the scatterplot as `stl_lead_plot`

-   Have the scatterplot returned in the knitted html - customize the size that it appears when knitted

-   Also save a .png of the scatterplot to `figs`, with dimensions of (6\" x 5\") (width x height)

-   In text above or below the scatterplot, write 1 - 2 sentences describing the overall trend that you observe from your graph

```{r}
stl_lead_plot <- ggplot(data = stl_lead_prop, 
                              aes(x = prop_white, y = pct_elevated))+
  geom_point(color = "red")+
  labs(title = "% with elevated blood lead levels vs % white in each census tract", 
       x = "pct. identifyin as white", 
       y = "elevated blood lead levels")+
  theme_minimal() 

ggsave(here::here("figs","ElevatedBloodLeadLevelsPct.png"),
       plot = stl_lead_plot, height = 5, width = 6)

stl_lead_plot

```

#### Observation 

The higher % of children identify as white, the lower % of elevated blood lead levels.

### **Create a histogram**

-   Create a histogram of only the `pctElevated` column in the data frame (remember, this will only take one variable - the frequency is calculated for you by `geom_histogram`)

-   Customize the fill, color, and size aesthetics - test some stuff! Feel free to make it awful.

-   Once you\'ve played around with customization, export the histogram as a .jpg to the `figs` folder

-   Make sure the histogram also shows up in your rendered html

```{r}
stl_lead_histo_plot <- ggplot(data = stl_lead_prop, 
                              aes(x = pct_elevated))+
  geom_histogram(fill = "red")+
  labs(title = "% of childred with elevated blood lead levels in each census tract", 
       x = "elevated blood lead levels")+
  theme_minimal() 

ggsave(here::here("figs","ElevatedBloodLeadLevelsHistogram.png"),
       plot = stl_lead_histo_plot, 
       height = 5, 
       width = 6)

stl_lead_histo_plot
```
